### 2015-02-05 cav-2015-tex/summary-heap.tex
minor wording changes and comments

\section{Summary Heap Framework}
\label{sec:sh}
The summary heap algorithm is defined using
Javalite~\cite{saints-MS}. The small-step semantics of Javalite,
defined as a syntactic machine using a CESK architecture, are modified
to perform symbolic execution on a summary heap. The surface (input)
and machine (state) syntax is shown
in~\figref{fig:syntax}. Terminals are in bold face while non-terminals
are italicized. Ellipses indicate zero or more repetitions. Tuples
omit the commas for compactness.

@@

A summary heap is a bipartite graph on references, $\cfgnt{r}$, and
constraint location pairs, $\lp\phi\ \cfgnt{l}\rp$. The graph is
expressed in the location map, $\cfgnt{L}$, and the reference map,
$\cfgnt{R}$. As done with the environment, $\cfgnt{L}$ and $\cfgnt{R}$
are treated as partial functions where $\cfgnt{L}(r) =
\{(\phi\ \cfgnt{l})\ ...\}$ is the set of constraint-location pairs in
the heap associated with the given reference, and
$\cfgnt{R}(\cfgnt{l},\cfgnt{f}) = \cfgnt{r}$ is the reference
associated with the given location-field pair in the heap. References
in the graph are immutable. Locations, however, are mutated on
updates. The structure with the immutability of references mean that
look-ups in the heap are non-recursive and updates do not modify the
current path condition.

The soundness and completeness proofs rely on a set of invariants that
are preserved by the small-step semantics during symbolic execution:
\emph{reference partitions}, \emph{null and uninitialized}, \emph{immutability}, \emph{determinism}, and
\emph{type consistency}.
\begin{compactdesc}
\item[Reference partitions] means that a reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap} that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in
  constraints); $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals} that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere); and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals} (these appear in environments, expressions, and continuations).
\item[Null and Uninitialized] means that every summary heap contains 
a special location for null ($\cfgnt{l}_\mathit{null}$), and a special
location for uninitialized
  ($\cfgnt{l}_\mathit{un}$) with $\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
  $\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$. 
\item[Immutability] means that a reference does not change in the
location map, and as a consequence, constraints never need to be
rewritten as the program evolves its state.
\item[Determinism] means that a reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
\item[Type consistency] means that all locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

@@

A key insight to the summary heap is that although many auxiliary or
stack literals are created in symbolic execution with lazy
initialization, only input heap references appear in constraints
expressing potential aliasing relationships. A solver reasons over the
aliasing relationships, and a satisfying assignment uniquely
identifies active edges in the graph. Those active edges describe a
representative concrete heap and its shape. For example, if the solver
assigns input heap variables $\cfgnt{r}$ and $\cfgnt{r}^\prime$ such
that $\cfgnt{r} = \cfgnt{r}^\prime$ is true, then there exists some
location $\cfgnt{l}$ and constraints $\phi$ and $\phi^\prime$ such
that (i) $\lp\phi\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r})$; (ii)
$\lp\phi^\prime\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r}^\prime)$; and
(iii) $\phi$ and $\phi^\prime$ hold in the aliasing assignments from
the solver.  In this way, the graph groups heaps with the conditions
under which those heaps exist so look-ups are non-recursive and
updates do not modify the path condition or require rewrites on
constraints.

>>>>

\section{Summary Heap Framework}
\label{sec:sh}

\sjp{summary heap or symbolic heap?}
The summary heap algorithm is defined using
Javalite~\cite{saints-MS}. The small-step semantics of Javalite,
defined as a syntactic machine using a CESK architecture, are modified
to perform symbolic execution on a summary heap. The surface (input)
syntax and machine (state) syntax are shown
in~\figref{fig:syntax}. Terminals are in bold face while non-terminals
are italicized. Ellipses indicate zero or more repetitions. Tuples
omit the commas for compactness.

@@

A summary heap is a bipartite graph on references, $\cfgnt{r}$, and
constraint location pairs, $\lp\phi\ \cfgnt{l}\rp$. The graph is
expressed in the location map, $\cfgnt{L}$, and the reference map,
$\cfgnt{R}$. As done with the environment, $\cfgnt{L}$ and $\cfgnt{R}$
are treated as partial functions where $\cfgnt{L}(r) =
\{(\phi\ \cfgnt{l})\ ...\}$ is the set of constraint-location pairs in
the heap associated with the given reference, and
$\cfgnt{R}(\cfgnt{l},\cfgnt{f}) = \cfgnt{r}$ is the reference
associated with the given location-field pair in the heap. References
in the graph are immutable. Locations, however, are mutated on
updates. 
%The structure with the 
Immutability of references means that
look-ups in the heap are non-recursive and updates do not modify the
current path condition.

The soundness and completeness proofs \sjp{in Section...} rely on a set of 
invariants \sjp{properties of the summary heap?} that
are preserved by the small-step semantics during symbolic execution:
%\emph{reference partitions}, \emph{null and uninitialized}, \emph{immutability}, \emph{determinism}, and
%\emph{type consistency}.
\begin{compactdesc}
\item[Reference partitions] means that a reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap} that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in
  constraints); $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals} that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere); and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals} (these appear in environments, expressions, and continuations).
\item[Null and Uninitialized] means that every summary heap contains 
a special location for null ($\cfgnt{l}_\mathit{null}$), and a special
location for uninitialized
  ($\cfgnt{l}_\mathit{un}$) with $\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
  $\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$. 
\item[Immutability] means that a reference does not change in the
location map, and as a consequence, constraints never need to be
rewritten as the program evolves its state.
\item[Determinism] means that a reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
\item[Type consistency] means that all locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

@@

\sjp{Up to now, no discussion of lazy initialization - what it is or how it is being used.}
A key insight to the summary heap is that although many auxiliary or
stack literals are created during symbolic execution with lazy
initialization, only input heap references appear in the constraints
expressing potential aliasing relationships. A solver is used to reason over the
aliasing relationships, and a satisfying assignment uniquely
identifies \emph{active edges} in the graph -- the set of edges describing a
representative concrete heap and its shape. For example, if the solver
assigns input heap variables $\cfgnt{r}$ and $\cfgnt{r}^\prime$ such
that $\cfgnt{r} = \cfgnt{r}^\prime$ is true, then there exists some
location $\cfgnt{l}$ and constraints $\phi$ and $\phi^\prime$ such
that (i) $\lp\phi\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r})$; (ii)
$\lp\phi^\prime\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r}^\prime)$; and
(iii) $\phi$ and $\phi^\prime$ hold in the aliasing assignments from
the solver.  In this way, the graph groups heaps with the conditions
under which those heaps exist so look-ups are non-recursive and
updates do not modify the path condition or require rewrites on
constraints.
